Journal and article coverage

### 2025-03-17 analysis/manuscript.qmd
Finish first full draft of result section

```{r country_data_prep}
jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
  filter(issn_l != "0027-8424")
oecd <- readr::read_csv("https://raw.githubusercontent.com/njahn82/hoa_ta_effects/refs/heads/main/data/oecd.csv")

@@

results are consistent.
most represented are chinese, usa and india in temrs of articles.
however, in terms of oa, other country lead,
these are in particular those with transformative agremments

overall, briscs and oecd, however, notable expection is south africa

en though correlation exisits, a closer lokk at some temproarl trend revela (dahsboard), which countries were affected by inconsitstent oa evidnecie.
heres is wiley, and there is france.

@@

@@

>>>>

When comparing countries, consistent patterns were observed across data sources from 2019 to 2023. 
Figure \ref{fig-uptake_country} presents hybrid open access indicators by country, comparing hoaddata (x-axis) with Web of Science and Scopus (y-axis).
Indicators calculated from these proprietary databases are shown for both first and corresponding authors, with full counting used to account for multiple country affiliations.

```{r country_data_prep}
#| fig-cap: "Open Access by country"
#| label: fig-uptake_country
#| 
jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
  filter(issn_l != "0027-8424")
oecd <- readr::read_csv("https://raw.githubusercontent.com/njahn82/hoa_ta_effects/refs/heads/main/data/oecd.csv")

@@

In terms of article output, a strong correlation was observed across the data sources.
Between 2019 and 2023, China emerged as the most productive country, followed by the United States and, by a wide margin, India, the United Kingdom, and Germany. 
Analysis of authorship roles revealed minimal variation, indicating that first and corresponding authors were typically from the same country.

When examining the percentage of open access articles in hybrid journals, a different pattern emerged.
Authors affiliated with institutions in medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access. 
Germany and the United Kingdom also had approximately 50% of their output available as open access.
In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the sample. 
The United States also demonstrated a relatively low proportion of open access articles.
These findings were consistent across all databases.
However, France was better represented in Web of Science, likely due to its agreement with Elsevier starting in 2019, which allowed delayed open access under the publisher’s user license [@Rabesandratana_2019]. 
This license was not classified as hybrid open access in Scopus or hoaddata.

Transformative agreements appeared to be a key driver of open access growth.
OECD members accounted for the majority of open access articles enabled by transformative agreements.
As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward [^sa].
Results were consistent across data sources, with a strong correlation observed. 
However, Wiley surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with Wiley that started in July 2019.

A closer examination of hybrid open access enabled by transformative agreements revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement. 
Medium-sized European countries again showed a high proportion of hybrid open access through transformative agreements, highlighting the impact of this licensing model across all three data sources.
In contrast, the United States had a low proportion of hybrid open access enabled by agreements, suggesting that a substantial number of open access articles were likely financed through other means, such as article processing charges (APCs).

@@

<!-- [juan study] larger proportion of hybrid, suggest journals included which might not be covered by transformative agreement. -->

@@

[^sa]: <https://sanlic.ac.za/read-and-publish-agreements/>

### 2025-03-17 analysis/manuscript.tex
Finish first full draft of result section

\section*{discussion}\label{discussion}
\addcontentsline{toc}{section}{discussion}

@@

>>>>

When comparing countries, consistent patterns were observed across data
sources from 2019 to 2023. Figure \ref{fig-uptake_country} presents
hybrid open access indicators by country, comparing hoaddata (x-axis)
with Web of Science and Scopus (y-axis). Indicators calculated from
these proprietary databases are shown for both first and corresponding
authors, with full counting used to account for multiple country
affiliations.

\begin{center}
\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/oa_country_comparision_plot_-1.pdf}
\end{center}

In terms of article output, a strong correlation was observed across the
data sources. Between 2019 and 2023, China emerged as the most
productive country, followed by the United States and, by a wide margin,
India, the United Kingdom, and Germany. Analysis of authorship roles
revealed minimal variation, indicating that first and corresponding
authors were typically from the same country.

When examining the percentage of open access articles in hybrid
journals, a different pattern emerged. Authors affiliated with
institutions in medium-sized European countries, such as Sweden, the
Netherlands, Finland, and Hungary, provided a large proportion of their
articles in open access. Germany and the United Kingdom also had
approximately 50\% of their output available as open access. In
contrast, non-OECD countries showed notably lower adoption of hybrid
open access, with South Africa being the only BRICS member
well-represented in the sample. The United States also demonstrated a
relatively low proportion of open access articles. These findings were
consistent across all databases. However, France was better represented
in Web of Science, likely due to its agreement with Elsevier starting in
2019, which allowed delayed open access under the publisher's user
license (Rabesandratana, 2019). This license was not classified as
hybrid open access in Scopus or hoaddata.

Transformative agreements appeared to be a key driver of open access
growth. OECD members accounted for the majority of open access articles
enabled by transformative agreements. As a notable exception, South
Africa also featured prominently, as the South African National Library
and Information Consortium (SANLiC) successfully negotiated
transformative agreements with major publishers from 2022 onward
\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}.
Results were consistent across data sources, with a strong correlation
observed. However, Wiley surplus in hoaddata led to better rankings for
countries where Wiley played a substantial role, such as Germany, where
the DEAL consortium negotiated its first transformative agreements with
Wiley that started in July 2019.

A closer examination of hybrid open access enabled by transformative
agreements revealed that the French Elsevier agreement on delayed open
access was not included in JCT data, as it did not qualify as a
transformative agreement. Medium-sized European countries again showed a
high proportion of hybrid open access through transformative agreements,
highlighting the impact of this licensing model across all three data
sources. In contrast, the United States had a low proportion of hybrid
open access enabled by agreements, suggesting that a substantial number
of open access articles were likely financed through other means, such
as article processing charges (APCs).

\section{discussion}\label{discussion}

@@

\bibitem[\citeproctext]{ref-Rabesandratana_2019}
Rabesandratana, T. (2019). Elsevier deal with france disappoints
open-access advocates. \emph{Science}.
\url{https://doi.org/10.1126/science.aba5656}
